---
title: "Getting started with the CANSIM package"
author: ""
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	collapse = TRUE,
	comment = "#>"
)
library(cansim)
```

# About Cansim

The `cansim` package provides R bindings to [Statistics Canada’s main socioeconomic time series database](https://www150.statcan.gc.ca/n1/en/type/data). Data can be accessed by table number, vector or table number and coordinate. The package accepts both old and new (NDM) cansim table numbers.

## Installing cansim

The stable version of **cansim** can be easily installed from CRAN.
```{r load_package_cran, echo=TRUE, message=FALSE, warning=FALSE, eval = FALSE}
install.packages("cansim")

library(cansim)
```

Alternatively, the latest development version can be installed from Github using `devtools`. 
```{r load_package_git, echo=TRUE, message=FALSE, warning=FALSE, eval = FALSE}
# install.packages("devtools")
devtools::install_github("mountainmath/cansim")

library(cansim)
```

## Usage

If you know the CANSIM table number you are interested in, use `get_cansim` to download the entire table.

```{r}
data <- get_cansim("14-10-0293")
head(data)
```

By defualt CANSIM data comes in the format provided by Statistics Canada, but often it is convenient to cast the data into an R data object and normalise values that are sent down in e.g. "millions" as unitless numbers. Piping data through the convenience function `normalize_cansim_values` takes care of that.
```{r}
data <- get_cansim("14-10-0293") %>% 
  normalize_cansim_values
head(data)
```

If you first want an overview of the data in the table you can use
```{r}
get_cansim_table_overview("14-10-0293")
```

If you don't know the table number you can browse the available tables or serach by survey, keyword or title.
```{r}
search_cansim_tables("housing price indexes")
```

Cansim data can also be accessed by vector, which is especially useful when building reports using specific indicators. For convenince one can specify named vectors, where the `label` field will be added to the returned tibble containing the specified name for each vector.
```{r}
get_cansim_vector(c("Meto Van Apartment Construction Price Index"="v44176267","Metro Van CPI"="v41692930"),start_time = "2015-05-01",end_time="2015-08-01")
```

